---
title: "Getting started with BSV"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with BSV}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

# BSV — Base of Security Vectors

`bsv` is an R-package that harvests cybersecurity research from arXiv,
enriches it with MITRE ATT&CK techniques and exposes the result through
a REST API (consumed by a Shiny dashboard and an MCP server for
AI agents).

The data flow is strictly:

```
arXiv API   ─►  ETL (R)   ─►  PostgreSQL   ─►  plumber2 API  ─►  Shiny / MCP
```

No component except the API touches the database directly. This makes
the data layer pluggable and the consumers easy to test.

## 1. Quick start (local R, no Docker)

You need a running PostgreSQL. The easiest way is `docker compose up db`,
then run R against it:

```{r}
library(bsv)

Sys.setenv(POSTGRES_HOST = "localhost",
           POSTGRES_USER = "bsv",
           POSTGRES_PASSWORD = "changeme",
           POSTGRES_DB = "bsv")

con <- db_connect()
db_init_schema(con)

df <- fetch_arxiv("cs.CR", max_results = 25)
df

save_articles(con, df)

res <- enrich_mitre(con, limit = 25, method = "rules")
res

db_disconnect(con)
```

## 2. Launching the API

In a separate R session:

```{r}
plumber::pr(system.file("api", "plumber.R", package = "bsv")) |>
  plumber::pr_run(host = "0.0.0.0", port = 8000)
```

Visit `http://localhost:8000/__docs__/` for the Swagger UI.

## 3. Launching the Shiny dashboard

```{r}
Sys.setenv(BSV_API_URL = "http://localhost:8000")
shiny::runApp(system.file("shiny", package = "bsv"))
```

## 4. Full stack via docker-compose

From the project root:

```bash
cp .env.example .env
docker compose up --build
```

This brings up four containers: `db`, `api`, `shiny`, `mcp`. Open
<http://localhost:3838> for the dashboard and POST JSON-RPC to
<http://localhost:8001/rpc> for the MCP endpoint.

## 5. Using LLM enrichment

By default `BSV_LLM_PROVIDER=none`, which uses the keyword-based rules in
`mitre_rules_dict()`. To switch to a real LLM, set in `.env`:

```bash
BSV_LLM_PROVIDER=openai
BSV_LLM_API_KEY=sk-...
BSV_LLM_MODEL=gpt-4o-mini
```

or, for a local Ollama:

```bash
BSV_LLM_PROVIDER=ollama
BSV_LLM_MODEL=llama3.1
BSV_OLLAMA_URL=http://ollama:11434
```

Then re-run `enrich_mitre(con, method = "llm")`. The HITL principle
applies: review the produced mappings in the dashboard and re-run on
the subset you disagree with.

## 6. MCP — talking to BSV from an AI agent

The MCP server speaks JSON-RPC 2.0 over HTTP. Available tools:

| Tool                  | Purpose                                       |
| --------------------- | --------------------------------------------- |
| `search_threat_intel` | Free-text search across title + abstract      |
| `list_articles`       | List with pagination                          |
| `get_article`         | One article + its MITRE mappings              |
| `mitre_stats`         | Aggregated counts per tactic / technique      |

Quick sanity check with `curl`:

```bash
curl -X POST http://localhost:8001/rpc -H "Content-Type: application/json" \
  -d '{"jsonrpc":"2.0","id":1,"method":"tools/list"}'
```
